Uncertainty propagation for predictor effects and bias correction in generalized linear (mixed) models

### 2023-02-22 paper/paper.tex
homing (台大)
ready to meet

The predicted \emph{direct} relationship between nitrogen and biomass is shown in the right panel. This does not track the visible relationship in this case, but instead reflects the modeled \emph{direct} effect of nitrogen. The meaning of such a plot will depend on which variables we choose to control for. Multivariate plots are not expected to be useful for prediction-style plots, so here we show only effect-style intervals.

@@

An example is shown in \fref{quadComp}. We fit a univariate, \emph{quadratic} response to nitrogen for the biomass. Both panels show the same model, and have the same response curves and prediction CIs. The panel on the left shows the default effect CIs, anchored at the model center. This does not correspond to any real value of nitrogen, because the mean of the quadratic values is taken separately, and is not the value that would correspond in that column to the mean of the original values. FOOTNOTE about Jensen, or what? \BC{Not at this point. Jensen has to do with the curvature of the link function. Here, we could simply say the model center does not correspond to a single input value but rather multiple predictor variables??} The panel on the right calculates effect CIs starting from the point that would correspond to an actual nitrogen value (in this case, the mean across the sampled values). The x-axis also reflects actual possible values of nitrogen; this is why the effect uncertainty goes to zero (the anchor point is visible) in the right panel but not the left panel. This is also why the response curves do not pass through the model-center point. 
\section{Generalized linear models}

>>>>

The predicted \emph{direct} relationship between nitrogen and biomass is shown in the right panel. This does not track the visible relationship in this case, but instead reflects the modeled \emph{direct} effect of nitrogen. The meaning of such a plot will depend on which variables we choose to control for. Prediction-style plots are less likely to be useful for multivariate models, so here we show only effect-style intervals.

@@

An example is shown in \fref{quadComp}. We fit a univariate, \emph{quadratic} response to nitrogen for the biomass. Both panels show the same model, and have the same response curves and prediction CIs. The panel on the left shows the default effect CIs, anchored at the model center. This does not correspond to any real value of nitrogen, because the mean of the quadratic values is taken separately, and is not the value that would correspond in that column to the mean of the original values. FOOTNOTE about Jensen, or what? \BC{Not at this point. Jensen has to do with the curvature of the link function.} \jd{Jensen is also how you show that the column means can never correspond to a real value of the parameter: the mean of the squares always exceeds the square of the mean.} \BC{Here, we could simply say the model center does not correspond to a single input value but rather multiple predictor variables??} The panel on the right calculates effect CIs starting from the point that would correspond to an actual nitrogen value (in this case, the mean across the sampled values). The x-axis also reflects actual possible values of nitrogen; this is why the effect uncertainty goes to zero (the anchor point is visible) in the right panel but not the left panel. This is also why the response curves do not pass through the model-center point.

In neither case is the picture as clear to interpret as the simplest case (\fref{splash}). We argue though that the picture on the left is a good way to show the shape of the model fit, while focusing on the uncertainty due to the parameter values due to the focal variable alone.

\section{Generalized linear models}
